---
title: "Hw13"
author: "Janeth Meraz"
date: "4/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Consider the large sample Z-test for the population proportion. Let p be the population proportion. The goal is to test the following hypothesis

H0:p=p0 vs. Ha:p/=p0.

```{r}
# function for the empirical level
emp_level = function(n = 40, R = 1000, p = 0.41, sigma = 1, alpha = 0.05) {
  
    p0 = p
    # store the test statistics
    stat = rep(NA, R)
    for (r in 1:R) {
        X = rbinom(n, 1, p0)
        stat[r] = (mean(X) - p0) / (sqrt(p0*(1-p0)/n))
    }

    # z critical value
    z_alpha = qnorm(1-(alpha/2))  
    cat(sprintf("Nominal level: %g\n", alpha))


    # Ha: p != p0
    emp_level = mean(abs(stat) > z_alpha)
    cat(sprintf("Ha: p /= p0, Empirical level: %g\n", emp_level))
    
    # Plot the histogram
    d = density(stat)  # kernel density
    hist(stat, xlab = "Test Statistic", prob = T, ylim = c(0, 
        max(d$y) + 0.05), main = "")

    # Add the null distribution and the kernel density on the
    # histogram.
    lines(d$x, d$y, col = "red")  
    curve(dt(x, df = n - 1), col = "blue", add = TRUE) 
    legend("topleft", legend = c("Kernel Density", "Null Distribution"), 
        lty = 1, col = c("red", "blue"), cex = 0.6)
}

# call empirical level
emp_level(n=40, R=1000, alpha=0.05)

# function for the power 
power_test = function(n = 40, R = 1000, p0 = 0.41, p = p0 + 
    0.40, sigma = 1, alpha = 0.05) {

    # store the test statistics
    stat = rep(NA, R) 
    for (i in 1:R) {
        X = rbinom(n, 1, p0)
        stat[i] = (mean(X) - p0) / (sqrt(p0*(1-p0)/n))
    }

    # Z_alpha: two sided critical  value
    Z_alpha = qnorm(1-(alpha/2))
    
    theoretical_power = 1 - pnorm(Z_alpha - (abs((p - p0) / (sqrt(p0*(1-p0)/n)))))
    cat(sprintf("Nominal level: %g, p = %g, theoretical power: %g\n", 
        alpha, p0, theoretical_power))

    # empirical power
    empirical_power = mean(abs(stat) < Z_alpha)
    cat(sprintf("Ha: p /= p0, Empirical power: %g\n", empirical_power))
    
    # plot the histogram
    d = density(stat) 
    hist(stat, xlab = "Test Statistic", prob = T, ylim = c(0, 
        max(d$y) + 0.05), main = "")

    # Add the null distribution and the kernel density on the
    # histogram.
    lines(d$x, d$y, col = "red")
    curve(dt(x, df = n - 1), col = "blue", add = TRUE) 
    legend("topleft", legend = c("Kernel Density", "Null Distribution"), 
        lty = 1, col = c("red", "blue"), cex = 0.6)
}

# show how the power of the test depends on alpha, p-p0 and n

power_test(n=20)
power_test(n=40)
power_test(n=50)

power_test(p = 0.5, p0 = 0.4)
power_test(p = 0.2, p0 = 0.9)
power_test(p = 0.44, p0 = 0.43)

power_test(alpha=0.05)
power_test(alpha=0.01)
power_test(alpha=0.20)

```

## Question 2

Consider the two-sample pooled t-test. Suppose X1,X2,...,Xn N(mu * x,sigma^2) and Y1,Y2,...,Ym N(mu * y,sigma^2) are independent random samples. We want to test the following hypothesis

H0:mux = muy vs. Ha:mux > muy.

Write a code for the empirical power function of this test.

Below, we have the code to show the empirical power of this test. 

```{r}
Power_Z_test = function(n = 40, m = 40, mux = 10, muy = 10, sd = 1, xbar = 50, ybar = 10, R = 1000, alpha = 0.05) {

    # Store the test statistics
    stat = rep(NA, R) 
    
    # Compute the test statistics from t distributions.
    for (i in 1:R) {
        X = rnorm(n, mux, sd)
        Y = rnorm(m, muy, sd)
        numerator = (mean(X)-mean(Y)) - (mux - muy)
        pooled = ((n-1)*var(X) + (m-1)*var(Y)) / (n+m-2)
        denominator = sqrt(pooled) * sqrt(1/n + 1/m)
        stat[i] = numerator/denominator
    }

    # T critical value
    t_alpha = qt(1 - alpha, n+m-2)

    # Empirical power
    empirical_power = mean(stat > -t_alpha)
    cat(sprintf("Ha: mux > muy, Empirical power: %g\n", empirical_power))
}

Power_Z_test()

```

